Suplementary Material 4: Draw Interbirth Interval Predictons on the Last Birth Omited Data

### 2023-04-18 S4_draw_IBI_predictions_EE_not_last.Rmd
replicate fig 4

---
title: 'Suplementary Material 3: Run Simulations on the Last Birth Omited Data'
author: "Richard Meitern"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
---

@@

The aim of this document is to display the R code needed the reproduce the simulations for the PISH scenarios. The letters P, I, S & H are used to indicate which mechanisms were included in each simulation scenario:

  * P indicates that a twinning event affects parity progression.
  * I indicates that a twinning event influences the time between that birth and the next one.
  * S indicates that a mother’s reproductive schedule affects both her likelihood of having twins and her total number of births, creating a link between these two factors.
  * H indicates that there is a connection between twinning propensity and intrinsic fertility due to maternal heterogeneity.

The scenario 0 does not include any of these mechanisms. This documents needs the twinR package to be installed to run.

@@

The memory (RAM) usage is quite high for simulations. The total number of CPUs used for running single simulation on Estonian data w/o last birth was 51. RAM usage varied between models but stayed around 18 GB per CPU core (dumped cores had a size range from 12 to 23GB).

```{r}
## Identify number of CPU cores available for parallel computing,
## note: using a large number may lead RAM to max out, so you may have to adjust
## that according to your infrastructure:
nb_cores <- 50

@@

## Running Simulations

```{r}
# import the function to do model fit and predictions
source("./R/fit_models.R")
```

The following is copied directly from twinR documentation and only the saving directories and
input data frames are changed.

```{r}
#------------------------------------------------------------------------------------------------
#-------------------------------- Goodness-of-fit tests -----------------------------------------
#------------------------------------------------------------------------------------------------

### IMPORTANT: while all the steps above should work no matter the operating system and whether
### you are using R within a GUI (e.g. RStudio) or not, the following step is very
### computationally intensive and has been tailored to be used on a Unix system (e.g. Linux) and
### directly within a terminal. It may work in RStudio, but this is not guaranteed, nor
### recommended. If you run this code using Windows, it should fallback to running the
### computation sequentially instead of in parallel across multiple CPU cores, which should work
### fine at the cost of requiring probably weeks of running time.

### Run scenarios one by one and save the output in a rda file:

library(twinR)
library(doSNOW)
#it seems that simulating the base model for Estonian not last birth data
#requires about 15GB of RAM per CPU core (tested with 61 CPU cores)

timeout <- 1 * 60 * 60 # an Hour

baseSlopeDir <- "./exports/slopes_under_scenarios"
baseFitDir <- "./exports/fitted_models"

scenarios_to_do <- c("base_model", "P", "I", "S", "H",
                     "PI", "PS", "PH", "IS", "IH", "SH",
                     "PIS", "PIH", "PSH", "ISH", "PISH")
```
```{r}
#use models without populations predictor
source("./R/twinR_models.R")
```

### TwinR data

```{r make directories}
expName <- "_EE_nl"
slopeDir <- paste0(baseSlopeDir, expName)
fitDir <- paste0(baseFitDir, expName)

dir.create(slopeDir, recursive = T)
dir.create(fitDir, recursive = T)

data_births_monthly <- data_births_monthly_EE_not_last

#remove unnecessary data frames
rm(data_births_monthly_EE, data_births_monthly_EE_not_last)

cat("Fits are saved in: ", fitDir)
```

```{r fit models}
fit_models  <- FALSE
if(fit_models){
  ## fit all trios of models in parallel (for linux only, but easy to adjust for other OS):
  pbmcapply::pbmclapply(scenarios_to_do, function(scenario) {
    targetFname <-  paste0(fitDir,"/fits_", scenario, "_obs.rda")
    if(!file.exists(targetFname)){
      fits <- fit_life_histories(scenario = scenario,
                               birth_level_data = data_births_monthly)
      save(fits, file =targetFname, compress = "xz")

      rm(fits)
    }

    }, mc.cores = min(c(length(scenarios_to_do), nb_cores)), mc.preschedule = FALSE)
}

@@

Its suggested to use a linux computing cluster to run the simulation scenarios one by one (using *run_simulation_EE_not_last.R*) rather than running the for loop below. Example of a bash command > **Rscript run_simulation_EE_not_last.R --model H --cores 51**. The model indicates the scenario to be run.

```{r run scenarios}
for (scenario in scenarios_to_do) {
   name_obj <- paste0("slopes_under_", scenario)
   targetFname <- paste0(slopeDir,"/", name_obj, ".rda")
   if(!file.exists(targetFname)){
     load(file = paste0(fitDir,"/fits_", scenario, "_obs.rda"))
     slopes_under_scenario <- simulate_slopes_for_GOF(N_replicates_level1 = 200L,
                                                      N_replicates_level2 = 20L,
                                                     birth_level_data = data_births_monthly,
                                                      life_history_fits = fits,
                                                      scenario = scenario,
                                                      nb_cores = nb_cores,
                                                      timeout = timeout,
                                                      .log = TRUE, lapply_pkg = "pbmcapply",
                                                      verbose = list(fit = TRUE, simu = FALSE))
     rm(fits)
     assign(name_obj, value = slopes_under_scenario)
     save(list = name_obj, file = targetFname)
     rm(list = name_obj) # remove the object behind the name!
     rm(slopes_under_scenario, scenario, name_obj) # remove the object directly
     gc_stuff <- gc(verbose = FALSE)
   }

}
```

```{r load slopes, include=FALSE}
## Combine simulated slopes:

all_slopes <- combine_simulated_slopes(path_slopes = slopeDir)
```

```{r}
#change the p-val expected direction
source("./R/twinR_goodness_of_fit.R")

# unchanged from the original but needed to have positive slopes yellow. 
source("./R/twinR_fig_5.R")
```

```{r draw figure}
## Figure 5 with colours inverted as we look for positive slopes:
fig5 <- draw_fig_5(all_slopes, width = 1)
fig5Fname <- paste0("./exports/fig5_", expName, ".")
ggplot2::ggsave(filename = paste0(fig5Fname, "pdf"), fig5,
                width = 88, height = 70, units = "mm")
fig5
```

```{r draw figures}
## test scenarios (table S13):

tableS13 <- goodness_of_fit(all_slopes)

writexl::write_xlsx(tableS13, paste0(fig5Fname, "xlsx"))
knitr::kable(tableS13)
```

```{r timing, results='asis'}
## computing time (for gof analysis):
timeHours <- computing_time_analysis(all_slopes)
cat("Total CPU time to run simulations was ",
    unlist(round(timeHours / 24, 2)), "days. ",
    "On a high RAM desktop computer with 8 cores it would hence take around",
    unlist(round(timeHours / 24 / 8, 0)), "days to run. However, on a standard ",
    "desktop computer with 24 GB RAM only 1 core could be used.")

>>>>

---
title: 'Suplementary Material 4: Draw Interbirth Interval Predictons on the Last Birth Omited Data'
author: "Richard Meitern"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
---

@@

The aim of this document is to display the R code needed the reproduce the original paper figure 4 equivalent on the Estonian dataset where the last births have been ommited.

@@

```{r}
## Identify number of CPU cores available for parallel computing,
## note: using a large number may lead RAM to max out, so you may have to adjust
## that according to your infrastructure:
nb_cores <- 7

@@

## Get Models

```{r}
fig_data_files <- sapply(LETTERS[1:3], function(l){
  predDir <- "./data/predictions/"
  paste0(predDir, "data_fig_4",l, ".rds")
})
```

```{r}
if(!all(sapply(fig_data_files, file.exists))){
  fitDir <- "D:/hpc/projects/natCom_twinR_commentary/exports/fitted_models_EE_nl/"
  load(paste0(fitDir, "fits_PISH_obs.rda")) # fit all

  data_fig_4A <- twinR::prepare_data_fig_4A(fits$fit_PP)
  gcs <- gc(verbose = F)
  saveRDS(data_fig_4A, paste0(predDir, "data_fig_4A", ".rds"))

  data_fig_4B <- twinR::prepare_data_fig_4B(fits$fit_IBI)
  gcs <- gc(verbose = F)
  saveRDS(data_fig_4B, paste0(predDir, "data_fig_4B", ".rds"))

  data_fig_4C <- twinR::prepare_data_fig_4C(fits$fit_twinning.binary)
  gcs <- gc(verbose = F)
  saveRDS(data_fig_4C, paste0(predDir, "data_fig_4C", ".rds"))
} else{
  for(l in LETTERS[1:3]){
    assign(paste0("data_fig_4",l), readRDS(fig_data_files[l]))
    }
}

@@

## Plot models

```{r}
fig_4A <- twinR::draw_fig_4A(data_fig_4A)
fig_4B <- twinR::draw_fig_4B(data_fig_4B)
fig_4C <- twinR::draw_fig_4C(data_fig_4C)
```

```{r}
twinR::plot_grid(fig_4A, fig_4B, fig_4C,
          labels = "auto", nrow = 1, label_size = 7, align = "v", axis = "lr")
```

```{r}
ggplot2::ggsave(file = "./exports/fig4.pdf", width = 180, height = 55, units = "mm")
```
